---
title: "Qitong’s Part"
author: "Qitong Liu"
date: "February 10, 2016"
output: pdf_document
---

```{r, echo=FALSE, message=FALSE}
############################
## EDAV PROJECT ONE
## THE CLASS
## 
## Shenghan Yu
############################

###########################################################################################

setwd("~/GitHub/ExploratoryDataAnalysis")
require(xlsx)
library(ggplot2)
library(dummies)

# import raw data
raw_data <- read.xlsx("Survey.xlsx", sheetName = "Form Responses 1")

# This is the dataset we use in the following 
dataset <- raw_data

###########################################################################################
## 1.Gender
dataset$gender[raw_data$What.is.your.preferred.gender.pronoun.=="he/him"] = "Male"
dataset$gender[raw_data$What.is.your.preferred.gender.pronoun.=="she/her"] = "Female"
dataset$gender[raw_data$What.is.your.preferred.gender.pronoun.=="doesn't matter"] = "Declined to state"
dataset$gender[is.na(raw_data$What.is.your.preferred.gender.pronoun.)] = "Declined to state"
dataset$gender <- factor(dataset$gender)

## 2.Program
dataset$program <- raw_data$Program
dataset$program[raw_data$Program == 'Ms in ds'] = 'IDSE (master)'
dataset$program[raw_data$Program == 'MSDS'] = 'IDSE (master)'
dataset$program[raw_data$Program == 'Data Science'] = 'IDSE (master)'
dataset$program[raw_data$Program == 'Applied Math'] = 'Other masters'
dataset$program[raw_data$Program == 'PhD Biomedical Informatics'] = 'Ph.D.'
dataset$program[raw_data$Program == 'QMSS'] = 'QMSS (master)'
dataset$program <- factor(dataset$program)

## 3.Waitlist
dataset$waitlist = factor(raw_data$Are.you.on.the.waiting.list.)

## 4.skill Dummies
# Get the skills for each samples.
skill_list <- strsplit(as.character(raw_data$Experiences.with.tools), ", ")

# Create a list for all skill names
skill_name <- vector(mode="character", length=0)

# Create dummies for skills, loop through samples and skills
for (sample_index in 1:nrow(raw_data)) {
  
  for (skill in skill_list[[sample_index]]) {
    
    # append the names, create a new column if it is a new skill
    if (!is.element(skill, skill_name)) {
      skill_name = append(skill_name,skill)
      dataset[,skill] = 0
    }
    
    dataset[sample_index,skill] = 1
  }
  
}

## 5 text editor
dataset$text_editor <- raw_data$What.code.text.editor.do.you.use.most.

## 6. Confidence
dataset$conf_r_manipulation <- factor(dataset$Programming.and.Analytical.Experiences..R..data.manipulation.and.modeling.,levels=c('None','A little', 'Confident', 'Expert'), ordered= TRUE) 
dataset$conf_r_graphic <- factor(dataset$Programming.and.Analytical.Experiences..R..graphic.basics..base..lattice..grid.etc....  ,levels=c('None','A little', 'Confident', 'Expert'), ordered= TRUE)
dataset$conf_r_multivariate <- factor(dataset$Programming.and.Analytical.Experiences..R..advanced..multivariate.data.analysis..e.g..spatiotemporal.data..visualization.and.modeling..    ,levels=c('None','A little', 'Confident', 'Expert'), ordered= TRUE)
dataset$conf_r_markdown <- factor(dataset$Programming.and.Analytical.Experiences..Reproducible.documentation.with.R..e.g..R.Markdown..  ,levels=c('None','A little', 'Confident', 'Expert'), ordered= TRUE)

dataset$conf_github <- factor(dataset$Programming.and.Analytical.Experiences..Github.,levels=c('None','A little', 'Confident', 'Expert'), ordered= TRUE)
dataset$conf_matlab <- factor(dataset$Programming.and.Analytical.Experiences..Matlab..data.manipulation..analysis..visualization.and.modeling.  ,levels=c('None','A little', 'Confident', 'Expert'), ordered= TRUE)

## Drop old columns
dataset <- dataset[,-(1:38)]

##################################################################################################

## Making subgroups

subgroup_R <- dataset[dataset$R==1,] 

subgroup_IDSE <- dataset[dataset$program=='IDSE (master)',] 
subgroup_nonIDSE <- dataset[dataset$program!='IDSE (master)',] 

subgroup_male <- dataset[dataset$gender=='Male',] 
subgroup_female <- dataset[dataset$gender=='Female',]
```


```{r,echo=FALSE, message=FALSE, warning=FALSE}
### word cloud of top 15 popular tools of the entire class
library(wordcloud)
words = colnames(dataset)[4:23]
words[9] = "Google Drive"
words[10] = "Regular Expressions"
words[12] = "Shell"
freq = colSums(dataset[,4:23])

# color sets to green
pal2 <- brewer.pal(9,"Greens")
wordcloud(words, freq, scale=c(4,.2), max.words = 15, random.order=FALSE, rot.per=0.1, colors=pal2, main="Title", use.r.layout=TRUE)
```

This graph shows the word cloud of top 15 popular tools of the entire class. Clearly, R, RStudio and Python are the most poplular ones.

```{r,echo=FALSE, message=FALSE, warning=FALSE}
### word cloud of top 10 popular tools between males (left) and females (right)
freq.m = colSums(subgroup_male[,4:23])
freq.f = colSums(subgroup_female[,4:23])
par(mfrow = c(1,2))
wordcloud(words, freq.m, scale=c(3,.2), max.words = 10, random.order=FALSE, rot.per=.1, colors=pal2, main="Title", use.r.layout=TRUE)
wordcloud(words, freq.f, scale=c(3,.2), max.words = 10, random.order=FALSE, rot.per=.1, colors=pal2, main="Title", use.r.layout=TRUE)
```

These two graphs shows the word cloud of top 10 popular tools between males (left) and females (right). We can see that Python, R, and  RStudio are still the most popular ones in males, while Python is less popular in females.

```{r,echo=FALSE, message=FALSE, warning=FALSE, fig.show="hide"}
### bubble chart of four R skills' confidence level
## for display reason, please use the pdf I saved
## install.packages("gridExtra")    # for lay out the six bubble charts
library(gridExtra)
conf_r = dataset[,25:28]
p = list()
ind = 1
for(i in 1:3){
  for(j in (i+1):4){
    conf_r_subset = as.data.frame(table(conf_r[,c(i,j)]))
    p[[ind]] = ggplot(conf_r_subset, aes_string(x=colnames(conf_r_subset)[1], y=colnames(conf_r_subset)[2], size=colnames(conf_r_subset)[3])) + geom_point() + scale_x_discrete(limits=c("None","A little","Confident","Expert")) + scale_y_discrete(limits=c("None","A little","Confident","Expert")) + scale_size_area(max_size = 10)
    ind = ind + 1
  }
}
do.call(grid.arrange,p)
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
## clean text_editor data
dataset$text_editor = as.character(dataset$text_editor)
dataset$text_editor[dataset$text_editor %in% c("sublime text 2", "sublime text", "Sublime Text 2", "Sublime Text!", "Sublime Text", "sublime", "Sublime 2")] = "Sublime"
dataset$text_editor[dataset$text_editor %in% c("textwrangler", "textWrangler")] = "Text Wrangler"
dataset$text_editor[dataset$text_editor %in% c("ipynb", "Ipython", "python", "IPython Notebook")] = "IPython"
dataset$text_editor[dataset$text_editor == "I used jupyter last semester"] = "jupyter"
dataset$text_editor[dataset$text_editor == "haven't used any"] = "None"
dataset$text_editor[dataset$text_editor %in% c("Atom/Sublime text", "Any (20 years C++/Java experience)", "Sublime Text / Eclipse", "Webstorm, pycharm")] = "Two or more"

### Histgram of text editors
ggplot(dataset, aes(x=reorder(text_editor, table(text_editor)[text_editor]))) + geom_bar() + coord_flip() + xlab("text editor") + ggtitle("Choices of Text Editors")
```

This bar graph clearly shows that RStudio is the dominant text editor that are used by most students in the class. The number of its users is over 4 times of the second text editor-Sublime.

```{r, echo=FALSE, message=FALSE, warning=FALSE}
### Histgram of gender and text editors
library(grid)
library(scales)
dataset_gender = dataset[dataset$gender != "Declined to state",]
perc.text.gender = as.data.frame(prop.table(table(dataset_gender$text_editor, dataset_gender$gender)))
perc.text.gender = perc.text.gender[perc.text.gender$Freq != 0,]
colnames(perc.text.gender) = c("text_editor", "gender", "percentage")
ggplot(perc.text.gender, aes(x=gender, y=percentage, fill=text_editor)) + geom_bar(position = "fill", stat = "identity") + scale_y_continuous(labels=percent_format()) + coord_flip() + theme(legend.position="bottom", legend.key.size=unit(0.3,"cm"), legend.text = element_text(size=8), legend.title = element_text(size=8, face="bold")) + guides(fill=guide_legend(nrow=2,byrow=TRUE)) + ggtitle("Choices of Text Editors within each Gender")
```

RStudio is still the most popular text editors in males and females. However, the percentages of males who use Sublime and Xcode are significant larger than females. The rest text editors are basically of the same percentages.